

% %NOTE:
% % CHECKED WITH SLIDES: YES!
% % CHECKED WITH EXERCISES: NO -- TODO
% % MISSING: Stuff about internet of things and trends -- but left out since not relevant

\section{Introduction}

\begin{definition}[Embedded System]
Information processing systems embedded into a larger product. Goal is not information processing! Information processing is the way to reach a certain goal.
\end{definition}

\begin{definition}[Hybrid systems]
The analog and digital system together.
\end{definition}

\begin{definition}[CPS]
cyber-physical system. When a system connects the physical world with the digital.
\end{definition}

\begin{definition}[Reactive system]
A reactive system is one which is in continual interaction 
with its environment and executes at a pace determined by 
that environment.
\end{definition}

\begin{definition}[Hard real-time constraint]
 real-time constraint is called hard, if not meeting that constraint could result in a catastrophe“
\end{definition}

\begin{definition}[Soft real-time constraint]
All non-hard real-time constraints.
\end{definition}

\begin{definition}[Guaranteed system response]
A guaranteed system response has to be explained without statistical arguments.
\end{definition}

\subsection{Where used?}
\begin{itemize}[noitemsep]
\item Logistics
\item Health Care
\item Building Automation
\item Factory Automation
\item Cars
\item Consumer Electronics (Microwave)
\end{itemize}



\subsection{Characteristics of ES}
\begin{itemize}[noitemsep]
\item dependable (zuverl\"assig)
\begin{itemize}[noitemsep]
\item Reliability: $R(t)$ is the probability of system working correctly provided that it was working at $t=0$
\item Maintainability: $M(d) $ is the probability of system working correctly d time units after error occurred
\item Availability: probability of system working at time t
\item Safety: no harm to be caused
\item Security: confidential and authentic communication
\end{itemize}
\item efficient
\begin{itemize}[noitemsep]
\item Energy efficient
\item Code-size efficient (especially for systems on a chip)
\item Run-time efficient
\item Weight efficient
\item Cost efficient
\end{itemize}
\item Dedicated towards a certain application. Makes no sense, to make the system faster than required
\item Dedicated user interface
\item real-time constraints: For real-time systems, right answers arriving too late are wrong.
\item A guaranteed system response has to be explained without statistical arguments
\end{itemize}


\subsection{ES vs. General Purpose System}

\begin{tabularx}{\textwidth}{ p{7cm}p{5cm} } %
Embedded Systems & General Purpose System \\ 

\cmidrule{1-2}

Few applications that are \newline
known at design-time & Broad class of applications.  \\
Not programmable by end 
user & Programmable by end user.  \\
Fixed run-time requirements 
(additional computing power 
not useful).  & Faster is better. \\ 
Criteria: 
\begin{itemize}[noitemsep]
\item cost
\item power consumption
\item predictability
\end{itemize}

&
Criteria:
\begin{itemize}[noitemsep]
\item cost
\item power consumption
\item average speed
\end{itemize}

\\

Hard and Soft deadlines, periods... &  Tasks do have timing constraints \\
Goal: Meet the individual timing requirements & Goal: Minimize the average response time, \newline without any guarantees \\

\cmidrule{1-2}



\end{tabularx}


\cleardoublepage